---
title: "ModelingPaper_GithubCode"
author: "Peter Arts"
date: "5/28/2024"
output: html_document
---

```{r setup, include = F}
library(tidyverse)
library(zoo)
library(patchwork)
library(scales)
library(deSolve)
library(fitdistrplus)

options(ggplot2.discrete.colour= MFluColors <- c('#00274C', #UMblue
                '#FFCB05', #Maize
                '#2F65A7', #Arboretum Blue
                '#9A3324', #Tappan Red
                '#D86018', #Ross Orange
                '#75988d', #Rackham Green
                '#A5A508', #Wave Field Green
                '#00B2A9', #Taubman Teal
                '#CFC096', #UMMA Tan
                '#702082', #A2 Amethyst
                '#575294' #Matthai Violet
                ))

options(ggplot2.discrete.fill= MFluColors <- c('#00274C', #UMblue
                '#FFCB05', #Maize
                '#2F65A7', #Arboretum Blue
                '#9A3324', #Tappan Red
                '#D86018', #Ross Orange
                '#75988d', #Rackham Green
                '#A5A508', #Wave Field Green
                '#00B2A9', #Taubman Teal
                '#CFC096', #UMMA Tan
                '#702082', #A2 Amethyst
                '#575294' #Matthai Violet
                ))

UMcolors <- c('#00274C', #UMblue
                '#FFCB05', #Maize
                '#2F65A7', #Arboretum Blue
                '#9A3324', #Tappan Red
                '#D86018', #Ross Orange
                '#75988d', #Rackham Green
                '#A5A508', #Wave Field Green
                '#00B2A9', #Taubman Teal
                '#CFC096', #UMMA Tan
                '#702082', #A2 Amethyst
                '#575294' #Matthai Violet
                )
```

```{r}
cohort1 <- read.table('Data/092722_allvirusdata_PSO.csv', header = T, sep = ',')
IDs <- sort(unique(cohort1$ID))
```

FITTING FECAL SHEDDING DATA
Gamma Sims: Fitting Fecal Shedding Data
```{r}
gamma_func=function(par,times){
  abs(par[3]) * dgamma(times, shape=abs(par[1]), scale = abs(par[2]), log = FALSE)
}

gamma_cost_func = function(par,data){

  k =abs(par[1])#shape
  theta= abs(par[2]) #scale
  mag = abs(par[3])
  t0= par[4] #offset

  times=data[,1]+t0
  conc=log10(data[,2])

  sim = gamma_func(c(k,theta,mag),times)

  cost = sum(abs(sim-conc)^2)+abs(t0)

}

gamma_sims <- list()

for (id in IDs){
  data <- subset(cohort1, cohort1[,1] == id & is.na(cohort1[,6]) == F)[,c(2,6,7)]
  out = optim(c(5, 2, 7, 0),
              gamma_cost_func,
              data=data,
              control=list(maxit=1E4))
  sim= cbind(-5:50 -out$par[4],
             gamma_func(out$par[1:3],-5:50))
  LOD = 50 #define limit of detection

  shedding_index = which(sim[,2]<8 & sim[,1] >= 0 & sim[,1] <= 30)
  shedding_range = range(sim[shedding_index,1])

gamma_sims[[as.character(id)]] <- data.frame('day_PSO'=sim[shedding_index,1], 'N_conc.' = sim[shedding_index,2])
}

gamma_df <- data.frame()
for (ID in names(gamma_sims)){
  gamma_df <- rbind(gamma_df,data.frame(ID,gamma_sims[[ID]]))
}
```

Weibull Sims: Fitting Fecal Shedding Data
```{r include=FALSE}
fit_func=function(par,times){
  # abs(par[3]) * dgamma(times, shape=abs(par[1]), scale = abs(par[2]), log = FALSE)
  # abs(par[3]) * dlnorm(times, meanlog = abs(par[1]), sdlog = abs(par[2]), log = FALSE)
  abs(par[3]) * dweibull(times, shape=abs(par[1]), scale = abs(par[2]), log = FALSE)
}

weibull_cost_func = function(par,data){

  k =abs(par[1])#shape
  theta= abs(par[2]) #scale
  mag = abs(par[3])
  t0= par[4] #offset

  times=data[,1]+t0
  conc=log10(data[,2])

  sim = fit_func(c(k,theta,mag),times)

  cost = sum(abs(sim-conc)^1.5)+0.5*abs(t0)^2 + abs(theta)^0.5 #10*log(abs(theta)) # + abs(theta)  + abs(k)
}


weibull_sims <- list()

for (id in IDs){
  data <- subset(cohort1, cohort1[,1] == id & is.na(cohort1[,6]) == F)[,c(2,6,7)]
  out = optim(c(1, 1, 1E2, 0),weibull_cost_func,data=data,control=list(maxit=1E4))
  sim= cbind(-5:50-out$par[4], fit_func(out$par[1:3],-5:50))
  LOD = 50 #define limit of detection

  shedding_index = which(sim[,2]<8 & sim[,1] >= 0 & sim[,1] <= 30)
  shedding_range = range(sim[shedding_index,1])

  weibull_sims[[as.character(id)]] <- data.frame('day_PSO'=sim[shedding_index,1], 'N_conc.' =sim[shedding_index,2])
}


weibull_df <- data.frame()
for (ID in names(weibull_sims)){
  weibull_df <- rbind(weibull_df,data.frame(ID,weibull_sims[[ID]]))
}
```

Use RMSE to determine better fit/ round day values
```{r}

gamma_df$day_PSO <- round(gamma_df$day_PSO)
weibull_df$day_PSO <- round(weibull_df$day_PSO)

shedrmse <- data.frame()

for (id in IDs){
  idsub <- cohort1[cohort1$ID == id,]

  if (nrow(idsub) >= 3 & max(idsub$Day)-min(idsub$Day) >= 10 & sum(idsub$N_det, na.rm =T) > 0) {
  gsub <- subset(gamma_df, gamma_df$ID == id)
  wsub <- subset(weibull_df, weibull_df$ID == id)
  csub <- subset(cohort1, cohort1$ID == id)

 grmse <- sqrt(mean(log10(csub$N_conc) - gsub$`N_conc.`[gsub$day_PSO %in% csub$Day], na.rm = T)^2)
 wrmse <- sqrt(mean(log10(csub$N_conc) - wsub$`N_conc.`[wsub$day_PSO %in% csub$Day], na.rm = T)^2)

 shedrmse <- rbind(shedrmse,data.frame('ID' = id,'Weibull' = wrmse, 'Gamma' = grmse))
}
}

shedrmse$gammabetter  <- (shedrmse$Weibull - shedrmse$Gamma > 0)

weibullIDs <- shedrmse$ID[shedrmse$gammabetter == F]
gammaIDs <- shedrmse$ID[shedrmse$gammabetter == T]


bestfit_sims <- data.frame()
for (id in as.character(weibullIDs)){
  bestfit_sims <- rbind(bestfit_sims,data.frame(id,weibull_sims[[id]]))
}
for (id in as.character(gammaIDs)){
 bestfit_sims <- rbind(bestfit_sims,data.frame(id,gamma_sims[[id]]))
}

bestfit_sims <- bestfit_sims[which(bestfit_sims$id != 4526 & bestfit_sims$id != 4546),]

##round day

bestfit_sims$day_PSO <- round(bestfit_sims$day_PSO,digits = 0)
colnames(bestfit_sims) <- c('ID', 'Day', 'N_conc')
bestfit_sims$N_conc <- 10^bestfit_sims$N_conc
bestfit_simsdf <- bestfit_sims

bestfit_sims <- bestfit_sims %>%
  split(~.$ID)

fitIDs <- names(gamma_sims)
threshold_sim_ids <- c()
for (id in fitIDs){
  IDsub <- subset(cohort1, cohort1$ID == id)
  if(nrow(IDsub) > 2 && (max(IDsub$Day - min(IDsub$Day))) > 14){
    threshold_sim_ids <- append(threshold_sim_ids, id)
  }
}

##what fraction of total individuals who met threshold did not have a single positive measurement
threshold_ids <- c()
threshold_neg_ids <- c()
for (id in IDs){
  IDsub <- subset(cohort1, cohort1$ID == id)
  if(nrow(IDsub) > 2 && (max(IDsub$Day - min(IDsub$Day))) > 14){
    threshold_ids <- append(threshold_ids, id)
    if(sum(IDsub$N_det, na.rm = T) == 0){
     threshold_neg_ids <- append(threshold_neg_ids, id)
    }
  }
}

threshold_pos_ids <- threshold_ids[! threshold_ids %in% threshold_neg_ids]

posprob <- 1- (length(threshold_neg_ids)/length(threshold_ids))

```

PLOT - Final Trajectories
``` {r}
# ggplot()+
#   geom_point(data = subset(cohort1,cohort1$N_det == T & cohort1$ID %in% bestfit_simsdf$ID), aes(x = Day, y = log10(N_conc)), shape = 19) +
#   geom_point(data = subset(cohort1,cohort1$N_det == F & cohort1$ID %in% bestfit_simsdf$ID), aes(x = Day, y = log10(N_conc)), shape = 1) +
#   geom_line(data = bestfit_simsdf,aes(x = Day, y = N_conc))+
#   facet_wrap(vars(ID))+
#   theme_bw()
# 
# ## 4526, 4546 should be removed
# 
```


New SARS Shedding Fits - New approach still couldnt give me a good model for 4526
```{r}
##what fraction of total individuals who met threshold did not have a single positive measurement
threshold_ids <- c()
threshold_neg_ids <- c()
for (id in IDs){
  IDsub <- subset(cohort1, cohort1$ID == id)
  if(nrow(IDsub) > 2 && (max(IDsub$Day - min(IDsub$Day))) > 14){
    threshold_ids <- append(threshold_ids, id)
    if(sum(IDsub$N_det, na.rm = T) == 0){
     threshold_neg_ids <- append(threshold_neg_ids, id)
    }
  }
}

threshold_pos_ids <- threshold_ids[! threshold_ids %in% threshold_neg_ids]

posprob <- 1- (length(threshold_neg_ids)/length(threshold_ids))
#_____________________________________________________

c1split <- cohort1 %>%
  split(~.$ID)

sarsthreshid <- c()
for (i in names(c1split)){
  idsub <- c1split[[as.character(i)]]
  
   if (nrow(idsub) >= 3 & max(idsub$Day) - min(idsub$Day) >= 10){
     sarsthreshid <- append(sarsthreshid,i)
   }
}
sarsthreshid

sarsthresh <- cohort1[which(cohort1$ID %in% sarsthreshid & is.na(cohort1$N_conc) == F),c('ID','Day','N_conc','N_det')]

sarsthresh.norm <- sarsthresh
##transform and normalize concentration data and day
sarsthresh.norm$N_conc <- log10(sarsthresh.norm$N_conc)

sarsthresh.norm$Day <- (sarsthresh.norm$Day-min(sarsthresh.norm$Day, na.rm = T))/
  max((sarsthresh.norm$Day-min(sarsthresh.norm$Day, na.rm = T)))

ggplot(sarsthresh.norm)+
  geom_point(aes(x = Day, y = N_conc, color = as.character(ID)))

cnormsplit <- sarsthresh.norm %>%
  split(~.$ID)

shedders <- c()
noshed <- c()
maxvec <- c()
for (i in names(cnormsplit)){
  
  maxvec <- append(maxvec, max(cnormsplit[[as.character(i)]]$N_conc))
  cnormsplit[[as.character(i)]]$N_conc <- cnormsplit[[as.character(i)]]$N_conc/max(cnormsplit[[as.character(i)]]$N_conc)
  cnormsplit[[as.character(i)]]$N_conc[cnormsplit[[as.character(i)]]$N_det == F] <- 0
  
  if(sum(cnormsplit[[as.character(i)]]$N_det) > 0 ){
    shedders <- append(shedders, i)
  } else {noshed <- append(noshed,i)}
}
names(maxvec) <- names(cnormsplit)

cnormsplit <- cnormsplit[names(cnormsplit) %in% shedders]

beta_cost_fun<- function(par, data){
  shape1 <- par[1]
  shape2 <- par[2]
  xscale <- par[3]
  yscale <- par[4]
  offset <- par[5]
  
  day <- append(0,data[,1]) ## Use 0,0 as data, how can I constrain this?
  conc <- append(0,data[,2])
  
  sim <- (dbeta(((day*xscale)+offset), shape1, shape2)/
            max(dbeta((day*xscale)+offset, shape1, shape2)))*abs(yscale)
  
  cost <- sqrt((sum((-sim+conc)^2))/length(conc))
  
}

ModIDs <- list('Beta' = c(4534,4547,4577,4579,4602,4520,4525,4532,4533,4538,4540,4541,4544,4545,4546,4549,4558,4576,4585,4586,4587,4591,4593,4596,4597,4598,4600))

beta_sims <- data.frame('ID' = as.character(ModIDs$Beta), 'Shape1' = NA, 'Shape2' = NA,'xScale' = NA, 'yScale' = NA, 'Offset' = NA,  'RMSE' = NA)
simdata <- list()

InitIDS <- list('A' = c('4525','4532','4533','4544','4545','4549','4584','4602'),
              'B' = c('4540','4541','4558','4597','4598','4600'),
              'C' = c('4544','4577','4585','4596','4593'),
              'D' = c('4591'),
              'E' = c('4546','4576','4587'),
              'F' = c('4520','4586'),
              'G' = c('4538'),
              'H' = c('4534'),
              'I' = c('4579'),
              'J' = c('4547'),
              'K' = c('4526'))

##Beta models
for (i in names(cnormsplit)){
  data <- cnormsplit[[as.character(i)]][,c(2,3)]
  
  if(i %in% names(cnormsplit)){
    for(j in names(InitIDS)){
      if (i %in% InitIDS[[j]]){ group <- j}
    }

    Inits <- list('A' = c(3,5,1,1,0.2-data$Day[data$N_conc == max(data$N_conc)]),
                  'B' = c(2,5,1,1,0.2-data$Day[data$N_conc == max(data$N_conc)]),
                  'C' = c(4,6,1,1,0.2-data$Day[data$N_conc == max(data$N_conc)]),
                  'D' = c(10,10,1,1,0.2-data$Day[data$N_conc == max(data$N_conc)]),
                  'E' = c(10,5,1,1,0.2-data$Day[data$N_conc == max(data$N_conc)]),
                  'F' = c(3,5,0.6,1,-0.1),
                  'G' = c(20,2,0.1,1,0),
                  'H' = c(7,2,2,1,-0.5),
                  'I' = c(2,5,0.3,1,0),
                  'J' = c(8,3,2,1,-0.4),
                  'K' = c(1,1,0.8,1,0.2-data$Day[data$N_conc == max(data$N_conc)]))
    
    par <- Inits[[group]]
    
    try(out <- optim(par,fn = beta_cost_fun, data = data,control=list(maxit=1E5)))
    
    beta_sims[beta_sims$ID == i,'Shape1'] <- out$par[1]
    beta_sims[beta_sims$ID == i,'Shape2'] <- out$par[2]
    beta_sims[beta_sims$ID == i,'xScale'] <- out$par[3]
    beta_sims[beta_sims$ID == i,'yScale'] <- out$par[4]
    beta_sims[beta_sims$ID == i, 'Offset'] <- out$par[5]
    beta_sims[beta_sims$ID == i,'RMSE'] <- out$value
    simdata[[as.character(i)]] <- out
  }
}
#renormalize data
predicts <- data.frame()
norm.predicts <- data.frame()

for (i in 1:nrow(beta_sims)){
  x.norm <- seq(0,5,length = 151)
  x <- ((x.norm)*30)-2
  
  c.norm <- dbeta((x.norm*beta_sims[i,'xScale'])+beta_sims[i,'Offset'], shape1 = beta_sims[i,'Shape1'], shape2 = beta_sims[i,'Shape2'])/
    max(dbeta((x.norm*beta_sims[i,'xScale'])+beta_sims[i,'Offset'], shape1 = beta_sims[i,'Shape1'], shape2 = beta_sims[i,'Shape2'])) *abs(beta_sims[i,'yScale'])
  
  c <- 10^((c.norm)*maxvec[beta_sims$ID[i]])
  
  predicts <- bind_rows(predicts, data.frame('ID' = beta_sims$ID[i],'Day'=x,'N_conc' = c))
  norm.predicts <- bind_rows(norm.predicts, data.frame('ID' = beta_sims$ID[i],'Day'=x.norm,'N_conc' = c.norm))
}

cnorm <- bind_rows(cnormsplit)


print(ggplot()+
  geom_line(data = norm.predicts,aes(x = Day, y = N_conc))+
  geom_point(data = cnorm,aes(x = Day, y = N_conc))+
  facet_wrap(vars(ID), scales = 'free_x')+
  theme_bw())

c <- bind_rows(predicts)

ggplot()+
  geom_line(data = c[c$N_conc > 1,],aes(x = Day, y = N_conc))+
  geom_point(data = sarsthresh[sarsthresh$ID %in% c$ID,],aes(x = Day, y = N_conc, shape = N_det), alpha = 0.5)+
  facet_wrap(vars(ID), scales = 'free_x')+
  scale_y_log10(labels = trans_format('log10',math_format(10^.x)), breaks = 10^seq(2,8, by =2), minor_breaks = 10^c(1:9))+
  scale_shape_manual(values = c(21,16))+
  guides(shape = 'none')+
  labs(y = 'gc/mL-wastewater', x = '')+
  theme_bw()

ggsave('Betafits.png',units = 'in', width = 10, height = 8)

predicts <- predicts %>%
  split(~.$ID)
```



BIOMARKER SHEDDING MODELS

PMMOV
```{r}
# pmmov <- cohort1$PMMoV_conc[is.na(cohort1$PMMoV_conc) == F]
# pmean <- mean(log10(pmmov))
# psd <- sd(log10(pmmov))
# 
# mean_pmmov_shed <- data.frame()
# for (id in IDs){
#   IDsub <- subset(cohort1, cohort1[,1] == id & is.na(cohort1$PMMoV_conc) == F)
#   indmean <- mean(log10(IDsub$PMMoV_conc))
#   indsd <- sd(log10(IDsub$PMMoV_conc))
#   mean_pmmov_shed <- bind_rows(mean_pmmov_shed, data.frame(id,indmean, indsd))
# }
# 
# clean_mean_pmmov_shed <- mean_pmmov_shed[is.na(mean_pmmov_shed$indsd) == F,]
# 
# pmmov_shedding <- function(logmean_pmmov, N, D){
#   pmmov_counter <- data.frame('Individual' = 1:N, 'pmmovID' = 1:N)
#   
#   pmmov_counter$pmmovID <- sample(1:nrow(logmean_pmmov), nrow(pmmov_counter), replace = T)
#   
#   shed_vect <- c()
#   for (day in 1:D){
#     
#     shed_vect <- rnorm(logmean_pmmov[pmmov_counter$pmmovID,2],
#           mean = logmean_pmmov[pmmov_counter$pmmovID,2], 
#           sd = logmean_pmmov[pmmov_counter$pmmovID,3])
#     
#     shed_vect[shed_vect > log10(max(cohort1$PMMoV_conc, na.rm = T))] <- log10(max(cohort1$PMMoV_conc, na.rm = T))
#     
#     pmmov_counter <- cbind(pmmov_counter,data.frame(shed_vect))
#   }
#   pmmov_counter
# }
```

beta distributed biomarkers
```{r}
c1split <- cohort1 %>%
  split(~.$ID)

sample_size <- map_dbl(c1split,nrow)

idthresh <- names(sample_size)[sample_size > 2]

dayrange <- c()
for (i in unique(cohort1$ID)){
  idsub <- cohort1[which(cohort1$ID == i),]
  dayrange <- append(dayrange ,max(idsub$Day, na.rm = T)
 - min(idsub$Day, na.rm = T))
}

names(dayrange) <- unique(cohort1$ID)

normtest <- data.frame()


CohThresh <- cohort1[cohort1$ID %in% names(sample_size >3),]

Cohnormp <- tibble('ID' = CohThresh$ID[which(CohThresh$PMMoV_det)], 
                   'Day' = CohThresh$Day[which(CohThresh$PMMoV_det)],
                  'PMMoV' = CohThresh$PMMoV_conc[which(CohThresh$PMMoV_det)])

Cohnormc <- tibble('ID' = CohThresh$ID[which(CohThresh$crAss_det)], 
                   'Day' = CohThresh$Day[which(CohThresh$crAss_det)],
                   'crAss' = CohThresh$crAss_conc[which(CohThresh$crAss_det)])


CohNorm <- full_join(Cohnormp, Cohnormc)
CohNorm$PMMoV <- log10(CohNorm$PMMoV)/log10(max(CohNorm$PMMoV, na.rm = T))
CohNorm$crAss <- log10(CohNorm$crAss)/log10(max(CohNorm$crAss, na.rm = T))

Cohsplit <- CohNorm %>%
  split(~.$ID)

pbeta <- data.frame()
cbeta <- data.frame()

pfit <- list()
cfit <- list()

pplot <- data.frame()
cplot <- data.frame()

CulFreyplotp <- ggplot()
CulFreyplotc <- ggplot()
for (i in idthresh){
  psub <- Cohsplit[[as.character(i)]]$PMMoV[which(is.na(Cohsplit[[i]]$PMMoV) == F)]
  csub <- Cohsplit[[as.character(i)]]$crAss[which(is.na(Cohsplit[[i]]$crAss) == F)]
  
  if (length(psub) >= 3){
    try(pfit <- fitdist(psub, distr = 'beta', method = 'mme'))
    pbeta <- bind_rows(pbeta, c('ID' = as.character(i),pfit$estimate[1], pfit$estimate[2]))
    dens <- dbeta(seq(0,1,length = 100), shape1 = pfit$estimate[1], shape2 = pfit$estimate[2])
    pplot <- bind_rows(pplot, data.frame('ID' = i, 
                                         'conc' = seq(0,1,length = 100),
                                         'density' = dens/max(dens)))
  }
  
  if (length(csub) >= 3){
    try(cfit <- fitdist(csub, distr = 'beta', method = 'mme'))
    cbeta <- bind_rows(cbeta, c('ID' = as.character(i), cfit$estimate[1], cfit$estimate[2])) 
    dens <- dbeta(seq(0,1,length = 100), shape1 = cfit$estimate[1], shape2 = cfit$estimate[2])
    cplot <- bind_rows(cplot, data.frame( 'ID' = i, 
                                          'conc' = seq(0,1,length = 100),
                                          'density' = dens/max(dens)))
  }
}

cplot$conc <- 10^(cplot$conc*log10(max(Cohnormc$crAss, na.rm = T)))

pplot$conc <- 10^(pplot$conc*log10(max(Cohnormp$PMMoV, na.rm = T)))

pplot <- pplot[which(is.infinite(pplot$density) == F),]
cplot <- cplot[which(is.infinite(cplot$density) == F),]


CohThresh <- cohort1[which(cohort1$ID %in% names(sample_size >3)),]

Cohnormp <- tibble('ID' = CohThresh$ID[CohThresh$PMMoV_det & is.na(CohThresh$PMMoV_det) == F], 
                   'Day' = CohThresh$Day[CohThresh$PMMoV_det & is.na(CohThresh$PMMoV_det) == F],
                  'PMMoV' = CohThresh$PMMoV_conc[CohThresh$PMMoV_det & is.na(CohThresh$PMMoV_det) == F])

Cohnormc <- tibble('ID' = CohThresh$ID[CohThresh$crAss_det & is.na(CohThresh$crAss_det) == F], 
                   'Day' = CohThresh$Day[CohThresh$crAss_det & is.na(CohThresh$crAss_det) == F],
                   'crAss' = CohThresh$crAss_conc[CohThresh$crAss_det & is.na(CohThresh$crAss_det) == F])


CohNorm <- full_join(Cohnormp, Cohnormc)
CohNorm$PMMoV <- log10(CohNorm$PMMoV)/log10(max(CohNorm$PMMoV, na.rm = T))
CohNorm$crAss <- log10(CohNorm$crAss)/log10(max(CohNorm$crAss, na.rm = T))

Cohsplit <- CohNorm %>%
  split(~.$ID)

pbeta <- data.frame()
cbeta <- data.frame()

pfit <- list()
cfit <- list()

pplot <- data.frame()
cplot <- data.frame()

CulFreyplotp <- ggplot()
CulFreyplotc <- ggplot()
for (i in unique(CohNorm$ID)){
  psub <- Cohsplit[[as.character(i)]]$PMMoV[is.na(Cohsplit[[as.character(i)]]$PMMoV) == F]
  csub <- Cohsplit[[as.character(i)]]$crAss[is.na(Cohsplit[[as.character(i)]]$crAss) == F]
  
  if (length(psub) >= 3){
    try(pfit <- fitdist(psub, distr = 'beta', method = 'mme'))
    pbeta <- bind_rows(pbeta, c('ID' = i,pfit$estimate[1], pfit$estimate[2]))
    dens <- dbeta(seq(0,1,length = 100), shape1 = pfit$estimate[1], shape2 = pfit$estimate[2])
    pplot <- bind_rows(pplot, data.frame('ID' = i, 
                                         'conc' = seq(0,1,length = 100),
                                         'density' = dens/max(dens)))
  }
  
  if (length(csub) >= 3){
    try(cfit <- fitdist(csub, distr = 'beta', method = 'mme'))
    cbeta <- bind_rows(cbeta, c('ID' = i, cfit$estimate[1], cfit$estimate[2])) 
    dens <- dbeta(seq(0,1,length = 100), shape1 = cfit$estimate[1], shape2 = cfit$estimate[2])
    cplot <- bind_rows(cplot, data.frame( 'ID' = i, 
                                          'conc' = seq(0,1,length = 100),
                                          'density' = dens/max(dens)))
  }
}

cplot$conc <- 10^(cplot$conc*log10(max(Cohnormc$crAss, na.rm = T)))

pplot$conc <- 10^(pplot$conc*log10(max(Cohnormp$PMMoV, na.rm = T)))

pplot <- pplot[which(is.infinite(pplot$density) == F),]
cplot <- cplot[which(is.infinite(cplot$density) == F),]



ggplot()+
  geom_histogram(data = Cohnormc,aes(x = crAss, y = after_stat(ndensity), color = 'crAss', fill = 'crAss'), alpha = 0.4)+
  geom_histogram(data = Cohnormp,aes(x = PMMoV, y = after_stat(ndensity), color = 'PMMoV', fill = 'PMMoV'), alpha = 0.5)+
  geom_line(data = cplot, aes(x = conc, y = density, color = 'crAss'))+
  geom_line(data = pplot, aes(x = conc, y = density, color = 'PMMoV'))+
  facet_wrap(vars(ID))+
  scale_x_log10(labels = trans_format('log10',math_format(10^.x)), breaks = 10^seq(2,8, by =2), minor_breaks = 10^c(1:9))+
  scale_color_manual(values = c('crAss' = '#00274C', 'PMMoV' = '#FFCB05'))+
  scale_fill_manual(values = c('crAss' = '#00274C', 'PMMoV' = '#FFCB05'))+
  labs(y = 'Normalized Density', x = 'Conc (copies/mL-WW)', fill = '')+
  guides(color = 'none')+
  theme_bw()


c1split <- cohort1 %>%
  split(~.$ID)

length(c1split[])/length(c1split)
Pfrac <- c()
Cfrac <- c()
for (i in names(c1split)){
  
  idsub <- c1split[[as.character(i)]]
  if(nrow(idsub) >=3){
    pp <- sum(idsub$PMMoV_det, na.rm = T) > 0
    cc <- sum(idsub$crAss_det,na.rm = T) > 0
    
    Pfrac <- append(Pfrac, pp)
    Cfrac <- append(Cfrac, cc)
  }
}

Pfrac <- sum(Pfrac)/length(Pfrac)
Cfrac <- sum(Cfrac)/length(Cfrac)



betabiomod <- function(pars, N, D, scale, fracshed){
  counter <- data.frame('Individual' = 1:N, 'ID' = 1:N)
  
  shedpop <- round(N*fracshed, digits = 0)
  
  counter$ID <- c(sample(1:(nrow(pars)), size = shedpop, replace = T), seq(0,0, length = N-shedpop))
  
  shed_vect <- c()
  for (day in 1:D){
    
    shed_vect <- rbeta(shedpop,
          shape1 = as.numeric(pars[counter$ID[1:shedpop],2]), 
          shape2 = as.numeric(pars[counter$ID[1:shedpop],3]))
    
    shed_vect <- c(10^(shed_vect*log10(scale)), seq(0,0, length = N-shedpop))
    
    counter <- cbind(counter,data.frame(shed_vect))
  }
  counter
}



```

crAssphage
```{r}
# mean_crass_shed <- data.frame()
# for (id in IDs){
#   IDsub <- subset(cohort1, cohort1[,1] == id & is.na(cohort1$crAss_conc) == F)
#   indmean <- mean(IDsub$crAss_conc)
#   indsd <- sd(IDsub$crAss_conc)
#   mean_crass_shed <- bind_rows(mean_crass_shed, data.frame(id,indmean, indsd))
# }
# 
# clean_mean_crass_shed <- mean_crass_shed[is.na(mean_crass_shed$indsd) == F,]
# 
# crass_shedding <- function(mean_crass, N, D){
#   
#   crass_counter <- data.frame('Individual' = 1:N, 'crAssID' = 1:N)
#   
#   crass_counter$crAssID <- sample(1:nrow(mean_crass), nrow(crass_counter), replace = T)
#   
#   
#   crass_counter
#   shed_vect <- c()
#   for (day in 1:D){
#     shed_vect <- c()
#     
#     shed_vect <- rnorm(mean_crass[crass_counter$crAssID,2],
#                        mean = mean_crass[crass_counter$crAssID,2], 
#                        sd = mean_crass[crass_counter$crAssID,3])
#     
#     shed_vect[shed_vect > max(cohort1$crAss_conc, na.rm = T)] <- max(cohort1$crAss_conc, na.rm = T)
#     shed_vect[shed_vect < 0] <- 0
#     
#     
#     crass_counter <- cbind(crass_counter,data.frame(shed_vect))
#     
#   }
#   crass_counter
# }

```

WW MODELING
Dilution into WW
``` {r}
total_dil <- function(Cf,Mf = 2.8E4,Uw = 1E5,Pshed = 1, Ptotal) { #Concentration of target - gc/ mg-dw, Mass rate of feces production (wet) - 28,000 mg-dry/day, solid fraction of feces - 0.2, daily water use - 100,000 mL/day.
  (Cf*Mf*Pshed)/(Uw*Ptotal)
  }
  
l_dil <- function(K,TSS = 0.28,Pshedders, Ptotal,ffecal,Cfeces){ #0.28 mg/mL - calculated from Rose + 100L/day water use
  
  (TSS*ffecal*Cfeces)/(Ptotal*(1+K*TSS))
}

s_dil <- function(K,TSS = 0.28,Pshedders, Ptotal,ffecal,Cfeces){
  
  (K*(TSS*ffecal*(Pshedders/Ptotal)*Cfeces))/(1+K*TSS)
}

```

SEIR MODELING
SEIR Model
```{r, include=FALSE}

dS <- function(B, S, I){
  -B*S*I
}

dE <- function(B, S, I, a){
  -B*S*I - a*E
}

dI <- function(a, E, y, I){
  a*E - y*I
}

dR <- function(y, I){
  y*I
}

SEIR.model<-function (init, beta.s, gamma.e, gamma.i, times) {
  library(deSolve)
  seir <- function(time, state, parameters) {
    with(as.list(c(state, parameters)), {
      dS <- -beta.s * S * I
      dE <- beta.s * S * I - gamma.e * E
      dI <- gamma.e * E - gamma.i * I
      dR <- gamma.i * I
      return(list(c(dS, dE, dI, dR)))
    })
  }
  parameters <- c(beta.s = beta.s, gamma.e = gamma.e, gamma.i = gamma.i)
  out <- ode(y = init, times = times, func = seir, parms = parameters)
  out <- as.data.frame(out)
  out$time <- NULL
  return(out)
}
```

Experiment
```{r}
N <- 10000
parms <- c(beta=0.2,gamma.1=0.1,gamma.e=0.2)
times <- 1:550
xstart <- c(S=N-1, E=1, I=0, R=0)

p = 1/N
init <- c(S = 1-p, E = 0, I = p, R = 0)
mod <- SEIR.model(init = init, beta.s = parms[1], gamma.e = parms[3], gamma.i = parms[2], times = times)

SEIR_out <- round(mod*N, digits = 0)

ggplot(mod)+
  #geom_path(aes(x = 1:nrow(mod), y = S), color = 'purple')+
  #geom_path(aes(x = 1:nrow(mod), y = E), color = 'green')+
  geom_path(aes(x = 1:nrow(mod), y = I), color = 'red')+
  #geom_path(aes(x = 1:nrow(mod), y = R), color = 'blue')+
  labs(y = 'Proportion', x = 'Time (days)')+
  theme_classic()

ggplot(SEIR_out)+
 # geom_path(aes(x = 1:nrow(mod), y = S), color = 'purple')+
  #geom_path(aes(x = 1:nrow(mod), y = E), color = 'chartreuse4')+
  geom_path(aes(x = 1:nrow(mod), y = I))+
  #geom_path(aes(x = 1:nrow(mod), y = R), color = 'cyan3')+
  labs(y = 'Individuals with \nActive Infection', x = 'Day of Outbreak')+
  theme_classic()

```



``` {r}
create_infectioncounter <- function(n){
  infectioncounter <- data.frame(1:n, vector(mode = 'numeric', length = n),vector(mode = 'numeric', length = n),vector(mode = 'numeric', length = n))
  colnames(infectioncounter) <- c('ID', 'Day_of_infection', 'Trajectory', 'traj_counter')
  infectioncounter
}

traj_shed_index <- function(traj,day,shed_fits){
  if (traj != 0 && is.na(shed_fits[[traj]][day,3]) == F){
    shed_fits[[traj]][day,3]
  } else { 0 }
}

add_one <- function(ori,day){
  if (ori >0){ day + 1 }
  else { 0 }
}

day_infection_counts_map <- function(I, R, shed_fits, n, d, infectioncounter, posprob){
  
  oldinfects <- nrow(subset(infectioncounter, infectioncounter[,2] != 0))
  totalinfects <- I+R
  
  if(oldinfects < totalinfects){
    newinfects <- totalinfects - oldinfects
    
    
    for (i in (oldinfects+1):(oldinfects+newinfects)){
      infectioncounter[i,3] <- rbinom(1,1,posprob)
      if (infectioncounter[i,3] == 1){
        infectioncounter[i,3] <- sample(1:length(names(shed_fits)), 1)
      }
      infectioncounter[i,2] <- d
    }
    
    
  }
  
  infectioncounter[[4]] <- map2_dbl(infectioncounter[[2]],infectioncounter[[4]], add_one)
  
  dayshed  <- map2_dbl(infectioncounter[[3]][],infectioncounter[[4]], traj_shed_index, bestfit_sims)
  
  infectioncounter <- infectioncounter %>%
    mutate(dayshed)
  
  names(infectioncounter)[names(infectioncounter) == 'dayshed'] <- d
  
  infectioncounter
}
```

SARS Dilution in WW
```{r}
begin <- Sys.time()
SARS_count_list <- list()
for (i in 1){
  SARS_count <- create_infectioncounter(n = N)
  begin1 <- Sys.time()
  
  SARS_count <- map2_dfc(SEIR_out[,3], SEIR_out[,4], day_infection_counts_map, shed_fits = bestfit_sims, n = N, d = )
  
  
  for (day in 1:nrow(SEIR_out)){
    
    SARS_count <- day_infection_counts_map(I = SEIR_out[day,3],
                                       R = SEIR_out[day,4],
                                       shed_fits =  bestfit_sims, 
                                       n = N, 
                                       d = day, 
                                       infectioncounter = SARS_count,
                                       posprob = posprob)
    
  }
  end1 <- Sys.time()
  p1.1 <- end1-begin1
  sums <- c()
  for (col in 5:ncol(SARS_count)){
    add <- sum(SARS_count[,col])
    sums <- append(sums, add)
  }
  
  incidence <- c()
  medianage <- c()
  medianshed <- c() 
  for (day in 1:nrow(SEIR_out)){
    dayincidence <- sum(SARS_count$Day_of_infection == day)
    incidence <- append(incidence, dayincidence)
    agevect <- day - subset(SARS_count$Day_of_infection, SARS_count$Day_of_infection <= day)
    agevect <- subset(agevect, agevect <= 30)
    shedsubvect <- SARS_count[[day+4]]
    
    medianage <- append(medianage, median(agevect))
    medianshed <- append(medianshed, median(shedsubvect))
  }
  
  SEIR_out$incidence <- incidence
  SEIR_out$median_infection_age <- medianage
  SEIR_out$median_shed_magnitude <- medianshed
  SEIR_out$incidence[1] <- 0
  
  ggplot()+
    geom_point(aes(x = 1:length(sums),y = sums))
  
  SARS_count_list[[i]] <- SARS_count
}

end <- Sys.time()
p1 <- end-begin

#Dilute fecal concentrations
ww_df <- data.frame()
begin2 <- Sys.time()
for (i in 1){
  
  SARS_tot_ww <- c()
  SARS_l_ww <- c()
  SARS_s_ww <- c()
  for (jour in 5:ncol(SARS_count_list[[i]])){
    
    totdil <- 0
    sdil <- 0
    ldil <- 0
    
    totdil <- sum(map_dbl(SARS_count[[jour]],total_dil,Ptotal = N))
    
    SARS_tot_ww <- append(SARS_tot_ww, totdil)
    # SARS_s_ww <- append(SARS_s_ww, sdil)
    # SARS_l_ww <- append(SARS_l_ww, ldil)
  }
  
  ww_iter <- data.frame('Day' = 1:length(SARS_tot_ww),'WW' = SARS_tot_ww,'Iter' = i)
  ww_df <- bind_rows(ww_df,ww_iter)
}

end <- Sys.time()
p2 <- end-begin2
p3 <- end- begin


ggplot(ww_df[])+
  geom_line(aes(x = Day, y = WW, color = as.character(Iter)))+
  facet_wrap(vars(Iter))+
  guides(color = 'none')+
  labs(x = 'Day', y = 'WW (gc/mL-total WW )')+
  theme_bw()
```

PMMoV Dilution in WW
```{r include=FALSE}

pmmov_counter <- pmmov_shedding(clean_mean_pmmov_shed, 1000, D = 100) #D = days

PMMoV_ww <- c()
for (jour in 3:102){
  PMMoV_day_ww <- 0
  ####Change
   PMMoV_day_ww <- total_dil(Kdp,TSS,Pshedders = 1, Ptotal = 1, ffecal = 1, Cfeces = mean(10^pmmov_counter[,jour]))
    
  
  PMMoV_ww <- append(PMMoV_ww, PMMoV_day_ww/Kdp)
}

PMMoV_rollmean <- rollmean(PMMoV_ww,7)

ggplot() +
  geom_point(aes(x = 1:length(PMMoV_ww), y = PMMoV_ww), alpha = 0.5)+
  geom_line(aes(x = 1:length(PMMoV_rollmean), y = PMMoV_rollmean), size = 1)+
  scale_y_continuous(name = 'PMMoV (cp/mL)')+
  theme_classic()+
  expand_limits(y = c(1,10000000))
```

CrAss Dilution in WW 
```{r eval=FALSE, include=FALSE}
crAss_ww <- c()

crAss_counter <- crass_shedding(clean_mean_crass_shed, N, 100)
for (jour in 3:102){

  
    crAss_day_ww <- total_dil(Kdp,TSS,Pshedders = 1, Ptotal = 1, ffecal = 1, Cfeces = mean(crAss_counter[,jour]))

  crAss_ww <- append(crAss_ww, crAss_day_ww/Kdp)
}

crAss_rollmean <- rollmean(crAss_ww,7)
```

EXPERIMENTS
Run the Model
```{r, include=FALSE}

ode(
  func=seirModel,
  y=xstart,
  times=times,
  parms=parms,
  method=c("ode45")
) %>% 
  as.data.frame() -> SEIR_out

SEIR_out <- round(SEIR_out, digits = 0)

ggplot(SEIR_out)+
  geom_path(aes(x = time, y = S/N), color = 'purple')+
  geom_path(aes(x = time, y = E/N), color = 'green')+
  geom_path(aes(x = time, y = I/N), color = 'red')+
  geom_path(aes(x = time, y = R/N), color = 'blue')+
  labs(y = 'Proportion', x = 'Time (days)')+
  theme_classic()


p = 1/N
init <- c(S = 1-p, E = 0, I = p, R = 0)
mod <- SEIR.model(init = init, beta.s = parms[1], gamma.e = parms[3], gamma.i = parms[2], times = 1:250)

ggplot(mod)+
  #geom_path(aes(x = 1:nrow(mod), y = S), color = 'purple')+
  #geom_path(aes(x = 1:nrow(mod), y = E), color = 'green')+
  geom_path(aes(x = 1:nrow(mod), y = I), color = 'red')+
  #geom_path(aes(x = 1:nrow(mod), y = R), color = 'blue')+
  labs(y = 'Proportion', x = 'Time (days)')+
  theme_classic()

SEIR_out <- round(mod*N, digits = 0)

ggplot(SEIR_out)+
 # geom_path(aes(x = 1:nrow(mod), y = S), color = 'purple')+
  #geom_path(aes(x = 1:nrow(mod), y = E), color = 'chartreuse4')+
  geom_path(aes(x = 1:nrow(mod), y = I), color = UMcolors[3])+
  #geom_path(aes(x = 1:nrow(mod), y = R), color = 'cyan3')+
  labs(y = 'Individuals with \nActive Infection', x = 'Day of Outbreak')+
  theme_classic()
```
Experiments
```{r}
#mean shedding

simple_SARS <- c()
for (inf in SEIR_out$I){
  simple_SARS_dil <- inf*total_dil(Cf = mean(cohort1$N_conc[cohort1$N_det == T], na.rm = T),Ptotal = N)
    
  simple_SARS <- append(simple_SARS,simple_SARS_dil)
}


#Random Shedding
rand_SARS <- c()
mod_Sars <- data.frame()
for (ind in names(bestfit_sims)){
  colnames(bestfit_sims[[ind]]) <- c('day','logsars')
  mod_Sars<- rbind(mod_Sars,bestfit_sims[[ind]])
}

rand_SARS<- c()
for (day in 1:nrow(SEIR_out)){
  dayshed <- c()
  for (inf in 1:SEIR_out$I[day]){
    (rand <- sample(mod_Sars[,3],1))
    dayshed <- append(dayshed,rand)
  }
  randmean <- sum(10^dayshed)
  rand_SARS_dil <- total_dil(Cf = randmean,Ptotal = N)
  rand_SARS <- append(rand_SARS,rand_SARS_dil)
}
rollmean_rand_SARS <- rollmean(rand_SARS, 7)
```

Figure 1
``` {r}
randplot <- ggplot()+
  geom_point(aes(x = 1:250,y = rand_SARS))+
  labs(x = 'Day',y = 'gc/mL-WW')+
  theme_bw()

meanplot <- ggplot()+
  geom_point(aes(x = 1:250,y = simple_SARS))+
  labs(x = 'Day',y = 'gc/mL-WW')+
  theme_bw()

mechplot <- ggplot(ww_iter)+
  geom_point(aes(x = Day,y = WW))

rand <- data.frame('Day' = 1:250, 'WW' = rand_SARS, 'app' = 'Random')
meanw <- data.frame('Day' = 1:250, 'WW' = simple_SARS,'app' = 'Mean')
mech <- data.frame('Day' = ww_iter$Day, 'WW' = ww_iter$WW, 'app' = 'Mechanistic')

fig1data <- bind_rows(meanw, rand) %>%
  bind_rows(mech)

ggplot(fig1data)+
  geom_path(aes(x = Day, y = WW, color = app))+
  labs(x = 'Day',y = 'gc/mL-WW')+
  scale_y_continuous(labels = comma)+
  facet_grid(rows = vars(app))+
  guides(color = 'none')+
  scale_color_manual(values = UMcolors[c(2,5,6)])+
  theme_bw()


ggsave('figure1.png', units = 'in', width = 5, height = 5)
```

Stackplot
``` {r}
c <- SARS_count_list[[1]]
colnames(c)[5:254] <- 1:250

c <- c %>%
  pivot_longer(5:254,names_to = 'Day', values_to = 'shed')

c$unique <- 1:250000

c$Day <- as.numeric(c$Day)
c$ID <- as.character(c$ID)


c$ww <- total_dil(Cf = c$shed,Ptotal = N)

c <- arrange(c, desc(ww))

ggplot(data = c)+
  geom_histogram(aes(x = Day, y = ww, group = desc(ww)),fill = 'white', color = 'black', linewidth = 0.1,stat = 'identity')+
  guides(fill = 'none')+
  theme_bw()

ggplot(data = c)+
  geom_area(aes(x = Day, y = ww, group = desc(ID)),fill = 'white', color = 'black', linewidth = 0.1,stat = 'identity')+
  guides(fill = 'none')+
  theme_bw()


```

Characterizing individual shedding data
```{r}


byday <- c %>%
  split(~.$Day)

for (i in names(byday)){
  byday[[i]]$frac <- byday[[i]]$ww/sum(byday[[i]]$ww)
}

d <- bind_rows(byday)

ggplot(data = d)+
  geom_histogram(aes(x = Day, y = frac, group = desc(frac)),fill = 'white', color = 'black', linewidth = 0.1,stat = 'identity')+
  guides(fill = 'none')+
  theme_bw()


byday <- d %>%
  split(~.$Day)


dayfracsum <- data.frame()
for (i in names(byday)){
  mn <- data.frame('median' = median(byday[[i]]$frac), 'mean' = mean(byday[[i]]$frac), 'sd' = sd(byday[[i]]$frac), 'max' = max(byday[[i]]$frac))
  dayfracsum <- bind_rows(dayfracsum,mn)
}

dayfracsum$I <- SEIR_out$I

```

Pop Iteration
```{r include=FALSE}
popiter <- c(seq(100, 900, by = 100))

spaghettiAll <- list()
SEIR_results <- list()
SARS_count_results <- list()

for (i in popiter){
  #SEIR model
  p = 1/i
    init <- c(S = 1-p, E = 0, I = p, R = 0)
    mod <- SEIR.model(init = init, beta.s = parms[1], gamma.e = parms[3], gamma.i = parms[2], times = 1:600)
    
    SEIR_out <- round(mod*i, digits = 0)
    
    SEIR_results[[as.character(i)]] <- SEIR_out[which(SEIR_out$I >= 1),]
    
}

for (j in 1:10){
for (i in popiter){
  print(c('pop' = i,'iter' = j))
  SEIR_out <- SEIR_results[[as.character(i)]]
  SARS_count <- create_infectioncounter(n = i)
    for (day in 1:nrow(SEIR_out)){
        
        SARS_count <- day_infection_counts_map(I = SEIR_out[day,3],
                                           R = SEIR_out[day,4],
                                           shed_fits =  bestfit_sims, 
                                           n = i, 
                                           d = day, 
                                           infectioncounter = SARS_count,
                                           posprob = 1- (length(threshold_neg_ids)/length(threshold_ids)))
        
    }
    
    means <- c()
      for (col in 5:(length(SEIR_out$I)+4)){
        avs <- sum(10^SARS_count[,col])
        means <- append(means, avs)
      }
      
      incidence <- c()
      for (day in 1:nrow(SEIR_out)){
        dayincidence <- sum(SARS_count$Day_of_infection == day)
        incidence <- append(incidence, dayincidence)
      }
      
      SEIR_out$incidence <- incidence
      
      SEIR_results[[as.character(i)]] <- SEIR_out
      SARS_count_results[[as.character(i)]] <- SARS_count

      
      SARS_ww <- c()

  
  SARS_tot_ww <- c()
  SARS_l_ww <- c()
  SARS_s_ww <- c()
  for (jour in 5:ncol(SARS_count_results[[as.character(i)]])){
    
    totdil <- 0
    sdil <- 0
    ldil <- 0
    
    totdil <- sum(map_dbl(SARS_count[[jour]],total_dil,Ptotal = N))
    
    SARS_tot_ww <- append(SARS_tot_ww, totdil)
    # SARS_s_ww <- append(SARS_s_ww, sdil)
    # SARS_l_ww <- append(SARS_l_ww, ldil)
  }
  
  ww_iter <- data.frame('Day' = 1:length(SARS_tot_ww),'WW' = SARS_tot_ww,'Iter' = j)
  
  if(j == 1){
    spaghettiAll[[as.character(i)]] <- ww_iter
  } else (spaghettiAll[[as.character(i)]] <- bind_rows(spaghettiAll[[as.character(i)]],ww_iter))
 
}
  
}


for (i in names(spaghettiAll)){
  spaghettiAll[[i]]$pop = as.numeric(i)
}

spaghettidf <- bind_rows(spaghettiAll)

ggplot(spaghettidf)+
  geom_point(aes(x = Day, y = WW), alpha = 0.01)+
  facet_wrap(vars(pop), scales = 'free')+
  guides(color = 'none')

```

```{r include = FALSE}

source('spagstitch.R')
regs <- list()

SEIR_results <- list()

spagGL <- spagGL %>%
  split(~.$pop)

for (i in as.numeric(names(spagGL))){
  #SEIR model
  p = 1/i
    init <- c(S = 1-p, E = 0, I = p, R = 0)
    mod <- SEIR.model(init = init, beta.s = parms[1], gamma.e = parms[3], gamma.i = parms[2], times = 1:600)
    
    SEIR_out <- round(mod*i, digits = 0)
    SEIR_out$pop <- as.numeric(i)
    SEIR_out$day <- 1:nrow(SEIR_out)
    SEIR_results[[as.character(i)]] <- SEIR_out[which(SEIR_out$I >= .05*max(SEIR_out$I)),]
    
}

#linear regression for each run, prevalence vs. daily concentration
for (popu in as.numeric(names(spagGL))){
  
  spagGL[[as.character(popu)]] <- spagGL[[as.character(popu)]][which(spagGL[[as.character(popu)]]$Day %in% SEIR_results[[as.character(popu)]]$day),]
  
  spagGL[[as.character(popu)]]$normday <- spagGL[[as.character(popu)]]$Day - (min(spagGL[[as.character(popu)]]$Day)-1)
  
  regs[[as.character(popu)]] <- data.frame()
  print(popu)
  
  SEIR_results[[as.character(popu)]]$normI <- SEIR_results[[as.character(popu)]]$I/popu
  
  for (m in unique(spagGL[[as.character(popu)]]$Iter)){
    ind_reg <- lm(subset(spagGL[[as.character(popu)]], 
                         spagGL[[as.character(popu)]]$Iter == m)$SARS_ww ~ (SEIR_results[[as.character(popu)]]$I))
    
    SARS_lin_pred <- predict(ind_reg)
    
    r_squared <- summary(ind_reg)$adj.r.squared
    slope <- ind_reg$coefficients[2]
    regs[[as.character(popu)]] <- rbind(regs[[as.character(popu)]],data.frame(r_squared, slope))
    
  }
}

##calculate average and sd of R^2 for each scale

stats <- data.frame()
for (popu in names(spagGL)){
  r_squared <- mean(regs[[as.character(popu)]]$r_squared)
  slope <- mean(regs[[as.character(popu)]]$slope)
  r_squaredsd <- sd(regs[[as.character(popu)]]$r_squared)
  slopesd <- sd(regs[[as.character(popu)]]$slope)
  r_squared_med <- median(regs[[as.character(popu)]]$r_squared)
  slope_med <-  median(regs[[as.character(popu)]]$slope)
  
  
  stats <- rbind(stats, data.frame(popu, r_squared,r_squaredsd, slope,slopesd, r_squared_med, slope_med))
}


regdf <- data.frame()
for (item in names(regs)){
  itdf <- regs[[item]]
  itdf$pop <- as.integer(item)
  regdf <- rbind(regdf,itdf)
}

regdf$pop <- as.numeric(regdf$pop)
stats$popu <- as.numeric(stats$popu)

ggplot(regdf)+
  geom_boxplot(aes(x = pop, y = r_squared, group = pop))+
  #geom_point(data = stats,aes(x = popu, y = r_squared_med), shape = '-', size = 10)+
  scale_x_log10()

ggplot()+
  geom_point(data = regdf[which(as.numeric(regdf$pop) >= 20),],aes(x = pop, y = slope*pop))+
  geom_point(data = stats[which(stats$popu >= 20),],aes(x = as.numeric(popu), y = slope_med*as.numeric(popu)), shape = '-', size = 10)+
  scale_x_log10()

ggplot(stats[which(stats$popu >= 20),])+
  geom_point(aes(x = popu, y = r_squared))+
  scale_x_log10()

ggplot(stats)+
  geom_point(aes(x = popu, y = r_squaredsd))+
  scale_x_log10()

ggplot(stats)+
  geom_point(aes(x = popu, y = slope))+
  scale_x_log10()+
  scale_y_log10()

ggplot(stats)+
  geom_point(aes(x = popu, y = slopesd))+
  scale_x_log10()+
  scale_y_log10()

##Regressions

est_fit <- lm(r_squared ~ log(popu), stats) ##This regression evaluates the change in fit suitability with increasing pop
est_var <- lm(r_squaredsd ~ log(popu), stats) ## THis regression evaluates the change in run to run variance in fit suitability with increasing pop

est_corr <- lm(log(slope) ~ log(popu), stats) # This regression evaluates the change in prev-WW correlation with increasing population
est_corrvar <- lm(log(slopesd) ~ log(popu), stats) #This regression evaluates the change in run to run variance in prev-ww correlation with increasing population

reg1 <- ggplot(stats)+
  geom_point(aes(x = popu, y = r_squared))+
  #geom_line(aes(x = popu, y = predict(est_fit)))+
  scale_x_log10()+
  theme_classic()+
  labs(x = 'Population', y = bquote('Mean - Adjusted '~R^2))

reg2 <- ggplot(stats)+
  geom_point(aes(x = popu, y = r_squaredsd))+
  #geom_line(aes(x = popu, y = predict(est_var)))+
  scale_x_log10()+
  theme_classic()+
  labs(x = 'Population', y = bquote('Standard Deviation - Adjusted'~R^2))

reg3 <- ggplot(stats)+
  geom_point(aes(x = popu, y = slope))+
  #geom_line(aes(x = popu, y = exp(predict(est_corr))))+
  scale_x_log10()+
  scale_y_log10()+
  theme_classic()+
  labs(x = 'Population', y = bquote('Mean - Slope'))

reg4 <- ggplot(stats)+
  geom_point(aes(x = popu, y = slopesd))+
  #geom_line(aes(x = popu, y = exp(predict(est_corrvar))))+
  scale_x_log10()+
  scale_y_log10()+
  theme_classic()+
  labs(x = 'Population', y = bquote('Standard Deviation - Slope'))

(reg1 +reg2)/ (reg3 + reg4)


maxes <- c()
for(item in names(spaghettiAll)){
  maxes <- append(maxes, max(spaghettiAll[[item]][,3]))
  
}
f <- c()
for (item in 1:length(N)){
  f <- append(f, (.15*N[item])/maxes[item])
}

```

```{r}
allspag <- bind_rows(spagGL)
allspag.norm <- allspag[allspag$I >= 0.01*as.numeric(allspag$pop),]

for (i in names(SEIR_results)){
  SEIR_results[[as.character(i)]]$Day <- 1:length(SEIR_results[[as.character(i)]]$I)
  SEIR_results[[as.character(i)]]$pop <- as.numeric(i)
}

SEIRdf <- bind_rows(SEIR_results)

plotlayout <- 
'AB
 AB
 AB
 CC
 CC'



ww_iter <- ggplot(allspag[which(as.numeric(allspag$pop) %in% c(100,1000,10000,100000) & allspag$Iter <=10),])+
  geom_point(aes(x = normday, y = SARS_ww, color = as.character(Iter)), size = 0.5 , alpha = 0.3)+
  facet_grid(rows = vars(pop), scales = 'free')+
  scale_y_continuous(labels = comma)+
  scale_color_discrete()+
  theme_bw()+
  guides(color = 'none')+
  #theme(legend.position = 'none', strip.text.y = element_blank()) +
  labs(title = 'a',x = 'Day',y = 'Concentrations (gc/mL)')

seir_iter <- ggplot(SEIRdf[which(SEIRdf$pop %in% c(100,1000,10000,100000)),])+
  geom_line(aes(x = Day, y = I))+
  facet_grid(rows = vars(pop), scales = 'free')+
  labs(title = 'b',y = 'Infected Individuals')+
  theme_bw()

reg_rsq <- ggplot(regdf[which(as.numeric(regdf$pop) >10),])+
  geom_point(aes(x = pop, y = r_squared), size = 0.5,alpha = 0.5)+
  geom_point(data = stats[which(stats$popu > 10),],aes(x = popu, y = r_squared_med), shape = '-', size = 10)+
  scale_x_log10(labels = comma)+
  labs(title = 'c',x = 'Sewershed Population', y = 'r^2')+
  theme_bw()

reg_slope <- ggplot(regdf[which(regdf$pop >10),])+
  geom_point(aes(x = pop, y = slope*pop), size = 0.5 ,alpha = 0.5)+
  geom_point(data = stats[which(stats$popu > 10),],aes(x = popu, y = slope_med*popu), shape = '-', size = 10)+
  scale_x_log10(labels = comma)+
  scale_y_continuous(labels = comma)+
  labs(title = 'd',x = 'Sewershed Population', y = 'Normalized Slope (gc/mL)')+
  theme_bw()

(ww_iter + seir_iter)/(reg_rsq + reg_slope)

ggsave('figure3.png', units = 'in', width =8, height = 7)

```

``` {r}
##Biomarker Population Iteration
N <- c(
  seq(10,90, by = 10), 
  seq(100, 900, by = 100),
  seq(1000,10000, by = 1E3),
  5E4,1E5)

TSS = .280
spaghettiPMMoV <- list()
spaghetticrAss <- list()
PMMoV_data <- list()
crAss_data <- list()
reps <- 1:10
day  <- 100
for (popu in N) {
  
  PMMoV_rollmean_app <- data.frame()
  crAss_rollmean_app <- data.frame()
  PMMoV_app <- data.frame()
  crAss_app <- data.frame()
  
  
  for (j in reps){
  
  pmmov_counter <- betabiomod(pars = pbeta, N = popu, D = day, scale = max(cohort1$PMMoV_conc, na.rm = T), fracshed = Pfrac)
  crAss_counter <- betabiomod(pars = cbeta, N = popu, D = day, scale = max(cohort1$crAss_conc, na.rm = T), fracshed = Cfrac)
  
  PMMoV_ww <- c()
  crAss_ww <- c()
  
  
  for (jour in 3:32){
    
    PMMoV_day_ww <- 0
    crAss_day_ww <- 0
    
      ##change
      PMMoV_day_ww <- total_dil(Pshed = popu,Ptotal = popu,  Cf = mean(pmmov_counter[,jour]))
      crAss_day_ww  <- total_dil(Pshed = popu,Ptotal = popu, Cf = mean(crAss_counter[,jour]))
      
    
    PMMoV_ww <- append(PMMoV_ww, PMMoV_day_ww)
    crAss_ww <- append(crAss_ww, crAss_day_ww)
    print(c("Iteration" = j,'Population' =  popu))
  }
  
  PMMoV_rollmean <- rollmean(PMMoV_ww,7)
  crAss_rollmean <- rollmean(crAss_ww,7)
      PMMoV_rollmean <- data.frame('Day' = 4:(length(PMMoV_ww)-3),'WW_rollmean' = rollmean(PMMoV_ww,7))
      crAss_rollmean <- data.frame('Day' = 4:(length(crAss_ww)-3), 'WW_rollmean' = rollmean(crAss_ww,7))
      PMMoV_ww <- data.frame('Day' = 1:length(PMMoV_ww), 'WW_conc' = PMMoV_ww)
      crAss_ww <- data.frame('Day' = 1:length(crAss_ww), 'WW_conc' = crAss_ww)
      
      PMMoV_rollmean <- data.frame('Iter' = rep(j, nrow(PMMoV_rollmean)), PMMoV_rollmean)
      crAss_rollmean <- data.frame('Iter' = rep(j, nrow(crAss_rollmean)), crAss_rollmean)
      PMMoV_ww <- data.frame('Iter' = rep(j, nrow(PMMoV_ww)), PMMoV_ww)
      crAss_ww <- data.frame('Iter' = rep(j, nrow(crAss_ww)), crAss_ww)
    
    PMMoV_rollmean_app <- rbind(PMMoV_rollmean_app, PMMoV_rollmean)
    crAss_rollmean_app <- rbind(crAss_rollmean_app, crAss_rollmean)
    PMMoV_app <- rbind(PMMoV_app, PMMoV_ww)
    crAss_app <- rbind(crAss_app, crAss_ww)
  }
  
  PMMoV_data[[as.character(popu)]] <- PMMoV_app
  crAss_data[[as.character(popu)]] <- crAss_app
  PMMoV_data[[as.character(popu)]]$pop <- popu
  crAss_data[[as.character(popu)]]$pop <- popu
  
  spaghettiPMMoV[[as.character(popu)]] <- PMMoV_rollmean_app
  spaghetticrAss[[as.character(popu)]] <- crAss_rollmean_app
  
  
}

bspag <- bind_rows(PMMoV_data)
cspag <- bind_rows(crAss_data)

biospag <- bspag

colnames(biospag)<- c('Iter','Day','PMMoV','Pop')

biospag$crAss <- cspag$WW_conc

biospag <- pivot_longer(biospag, cols = c(PMMoV,crAss), names_to = 'target')

pop.labs <- c('100'= 100, '1,000'=1000, '10,000' = 10000,'100,000' = 1e+05)

scientific <- function(x){
    ifelse(x==0, "0", parse(text=gsub("[+]", "", gsub("e", " %*% 10^", scientific_format()(x)))))
}

#biomarker stats

biostats <- data.frame()
popmean <- data.frame()
biodata <- data.frame()
bstats <- data.frame()
cstats <- data.frame()

iter <- 1
for(pop in N){
  PMMoV <- biospag$value[which(biospag$Pop == pop & as.numeric(biospag$Iter) == iter & biospag$target == 'PMMoV')]
  
  crAss <- biospag$value[which(biospag$Pop == pop & biospag$Iter == iter & biospag$target == 'crAss')]
    
   
  biodata <- rbind(biodata, data.frame(pop, PMMoV, crAss))
  bpmean <- c()
  bpsd <- c()
  bcmean <- c()
  bcsd <- c()
  itvec <- c()
  for(iter in 2:10){
    bpmean <- append(bpmean,mean(biospag$value[biospag$Pop == pop & biospag$Iter == iter & biospag$target == 'PMMoV']))
    bpsd <- append(bpsd,sd(biospag$value[biospag$Pop == pop & biospag$Iter == iter& biospag$target == 'PMMoV']))
    bcmean <- append(bcmean,mean(biospag$value[biospag$Pop == pop & biospag$Iter == iter & biospag$target == 'crAss']))
    bcsd <- append(bcsd,sd(biospag$value[biospag$Pop == pop & biospag$Iter == iter& biospag$target == 'crAss']))
   
    itvec <- append(itvec,iter)
  }
  bstats <- bind_rows(bstats, data.frame('pop' = pop,'iter' = itvec,'target' = 'PMMoV', 'mean' = bpmean,'sd' = bpsd))
    cstats <- bind_rows(cstats, data.frame('pop' = pop,'iter' = itvec,'target' = 'crAss', 'mean' = bcmean,'sd' = bcsd) )
}

biostats <- bind_rows(bstats,cstats)

c <-biostats %>%
  split(~.$pop+.$target)

runtorun <- data.frame()
for (i in names(c)){
  sd <- sd(c[[as.character(i)]]$mean)
  mean <- mean(c[[as.character(i)]]$mean)
  
  runtorun <- bind_rows( runtorun,data.frame('pop' = unique(c[[as.character(i)]]$pop),'target' =  unique(c[[as.character(i)]]$target), 'sd' = sd, 'mean' = mean))
}



biocols  <- c('PMMoV' = "#E7B800","crAss" = "steelblue")

biodata <- pivot_longer(biodata, cols = 2:3, names_to = 'target', values_to = 'conc')

biodata$conc <- as.numeric(biodata$conc)

b<- ggplot(biodata[which(biodata$pop %in% c(100,1000,10000,100000)),])+
  geom_boxplot(aes(y = conc, color = target, fill = target), alpha = 0.3)+
  theme_classic()+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  facet_grid(cols = vars(pop))+
  #scale_y_continuous(labels = scientific, breaks = seq(1e6, 8e6, by = 1e6))+
  #expand_limits(y = c(0,6e6))+
  guides(color = 'none', fill = 'none')+
  labs(title = '',x = '', y = 'Concentration (gc/mL)', color = '', fill = '')



c<- ggplot(biostats, aes(x = pop, y = mean, color = target)) +
  geom_point(alpha = 0.5)+
  #geom_boxplot(data = biostats[which(biostats$target == 'PMMoV'),], aes(x = pop, y = mean, group = pop), color = '#E7B800', alpha = 0.4, outlier.alpha = 0, position = 'dodge')+
  #geom_boxplot(data = biostats[which(biostats$target == 'crAss'),],aes(x = pop, y = mean, group = pop), color = 'steel blue',alpha = 0.4, outlier.alpha = 0)+
  scale_x_log10(label = comma, breaks = c(10,100,1000,10000,100000) , minor_breaks = biostats$pop)+
  #scale_y_continuous(breaks = seq(0,9E7, by = 2E7), minor_breaks = seq(0,9E6, by = 5E6), label = comma)+
  #stat_smooth()+
  labs(, x = 'Population', color = '',y = 'Mean')+
  theme_bw()

d <- ggplot(biostats, aes(x = pop, y = sd/mean, color = target)) +
  geom_point(alpha = 0.5)+
  scale_x_log10(label = comma)+
  #scale_y_continuous(breaks = seq(0,5,by = 1), minor_breaks = seq(0,5, by = 0.25))+
  #stat_smooth()+
  labs(y = 'CV', x = 'Population', color = '')+
  theme_bw()

e <- ggplot(runtorun, aes(x = pop, y = (sd/mean), color = target)) +
  geom_point(alpha = 0.5)+
  scale_x_log10(label = comma)+
  #scale_y_continuous(breaks = seq(0,5,by = 1), minor_breaks = seq(0,5, by = 0.25))+
  #stat_smooth(method = 'lm')+
  labs(y = 'Pooled CV', x = 'Population', color = '')+
  theme_bw()

a<- ggplot(biospag[which(biospag$Pop %in% c(100,100000) & biospag$Iter == 1),]) +
  geom_path(aes(x = Day, y = value, color = target), alpha = 0.4)+
  geom_point(aes(x = Day, y = value, color = target), alpha = 0.5)+
  facet_grid(rows = vars(Pop))+
  labs(x = 'Day', y = 'Concentration (gc/mL)', color = '')+
  #scale_y_continuous(labels = scientific, breaks = seq(0, 2e7, by = 2e6))+
  expand_limits(y = c(1e5,1e7))+
  theme_bw()+
  theme(legend.position = 'right',strip.background = element_rect(fill = 'white'))

(c+d)/(e+a) + plot_layout(guides = 'collect')

ggsave('figure4.png', units = 'in', width = 10, height = 7)
```


